---
title: "Mixed model - Winter's politeness data"
author: "Julia Müller"
date: "12 10 2020"
output: html_document
---

# Packages
```{r}
library(tidyverse)
library(lme4)
library(afex)
library(MuMIn) 
library(lattice)
```

# Data and initial questions
```{r}
politeness <- read.csv("data/politeness_data.csv", stringsAsFactors = TRUE)
```

frequency: voice pitch in Hertz
attitude: polite or informal
scenario = item (e.g. one scenario was "asking for a favour")

The model should predict frequency based on attitude and gender.

Which variables should be treated as random effects?
Why can't we just fit a regression instead of a mixed model?

## Visualisations
Create boxplots of
(a) the frequency for each subject
(b) the frequency for each scenario
```{r}


```

# Factor coding
Make sure your categorical factors are coded as factors. Make a choice about factor coding and code it explicitly.
```{r}

```


# Model fitting
Construct a full model (random intercepts and random slopes). What do you notice? 

Then fit a reduced model.
```{r}

```

Interpretation of fixed effects:

## Model comparisons

### Fixed effects
Create a null model (intercept only), one model with only gender as a predictor, and compare the two to the full model.
```{r}

```

Which model has the best fit?

### Random effects
Let's check if both random effects seem necessary.
To compare the random effects structure of models, we need to set REML = TRUE in the model call and refit = FALSE in the anova call.
Model without subject intercept has singular fit, again.

```{r}

```
Which model has the best fit?


## Bonus: Model fitting with afex
Winter uses the afex package, try this out instead of lme4 and compare the output. Hint: the call starts with mixed()
```{r}

```

### R squared
How can you calculate the R2 for a mixed model? 
```{r}

```

### Bonus: Lattice package "Caterpillar plots"
Check out the documentation for the lattice package. Adapt the following code for your winner model. Can you interpret what this is plotting? Hint: Replace "id" with a term that has both a random intercept and a random slope.
    dotplot(ranef(mdl_fortp),
    scales = list(x = list(relation = 'free')))[['id']]
    
```{r}

```
  
