---
title: "Regression Review"
author: "Kyla McConnell & Julia Müller"
date: "8/31/2020"
output: html_document
---

## Regression review

In this exercise, we'll use a Tidy Tuesday dataset to get back into regression models to help us understand mixed effects models, which we'll get into within the next few sessions. 
First, take a look at the dataset (more info [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md)).

```{r}
library(tidyverse)
library(tidymodels)
library(corrplot)
library(car)
library(dotwhisker)
library(rms)

coffee <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')
```


### Data wrangling

Starting off with a small data wrangling challenge: 
a) Only keep the variables total_cup_points, species, country_of_origin, processing_method, aroma, flavor, aftertaste, acidity, body, balance, uniformity, clean_cup, sweetness, and cupper_points
b) Convert species and country_of_origin to factors
c) Save the resulting data frame as "coffee".
d) Additionally, create a new dataset called "arabica" which only contains arabica beans.

```{r}

```


### Modeling

Using this data, we'd like to answer the following questions:
(1) For the arabica beans only: Do ratings for aroma, flavor, balance, and clean_cup predict the overall rating (total_cup_points) of the coffee?
(2) Can we use the variables acidity, sweetness, clean_cup, and balance to predict the type of bean?

For one of the hypotheses above, think about how you could use a regression model to answer the question. You may have to wrangle a bit more first. 

First, do some exploratory data analysis about your question. This may include some descriptive statistics, some graphing, (tidyverse) summarizing, checking out correlations, etc.:

```{r}

```

Then prepare your predictors and fit the most appropriate type of model: 

```{r}

```

And check the assumptions:
```{r}

```

Check model fit. If appropriate, use predict() to see what the model predicts for different values: 
```{r}

```


### Graphing

For your hypothesis above, graph the model output with a dot-whisker plot or other appropriate plot style. Have a look at the dotwhisker package (vignette [here](https://cran.r-project.org/web/packages/dotwhisker/vignettes/dotwhisker-vignette.html)) and Ch 11 in Winter 2020:

```{r}

```


### Reporting

Now imagine that you want to present your work above. Write your analysis using precise and accurate statistical terminology. 

"

...


"


If you have time, make a nice markdown file with the important output, graphs and interpretation (without showing any unnecessary code.)