---
title: "Regression Review"
author: "Kyla McConnell & Julia Müller"
date: "8/31/2020"
output: html_document
---

## Regression review

In this exercise, we'll use a Tidy Tuesday dataset to get back into regression models to help us understand mixed effects models, which we'll get into within the next few sessions. 

First, take a look at the dataset (more info here: https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md)

```{r, error=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(corrplot)
library(car)
library(dotwhisker)
library(rms)

coffee <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')
```

### Data wrangling

Starting off with a small data wrangling challenge: 
a) Only keep the variables total_cup_points, species, country_of_origin, processing_method, aroma, flavor, aftertaste, acidity, body, balance, uniformity, clean_cup, sweetness, and cupper_points
b) Convert species and country_of_origin to factors
c) Save the resulting data frame as "coffee".
d) Additionally, create a new dataset called "arabica" which only contains arabica beans.

```{r}
coffee <- coffee %>% 
  select(c(total_cup_points, species, country_of_origin, processing_method, aroma, flavor, aftertaste, acidity, body, balance, uniformity, clean_cup, sweetness, cupper_points)) %>% 
  mutate(species = factor(species), 
         country = factor(country_of_origin))

arabica <- coffee %>%
  filter(species == "Arabica")
```

### Modeling

Using this data, we'd like to answer the following questions:
(1) For the arabica beans only: Do ratings for aroma, flavour, balance, and clean_cup predict the overall rating (total_cup_points) of the coffee?
(2) Can we use the variables acidity, sweetness, clean_cup, and balance to predict the type of bean?

For one of the hypotheses above, think about how you could use a regression model to answer the question. You may have to wrangle a bit more first. 

First, do some exploratory data analysis about your question. This may include some descriptive statistics, some graphing, (tidyverse) summarizing, checking out correlations, etc.:

#### Question 1

```{r, error=FALSE, warning=FALSE}
#Look at means of each group (rounded to the 10s for a better overview)
arabica %>% 
  group_by(round(total_cup_points/10, 0)) %>% 
  summarise(mean(aroma), mean(flavor), mean(balance), mean(clean_cup))

#Plot line graphs of predictors vs. DV
ggplot(aes(x=total_cup_points, y=aroma), data=arabica) +
  geom_smooth()

ggplot(aes(x=total_cup_points, y=flavor), data=arabica) +
  geom_smooth()

ggplot(aes(x=total_cup_points, y=balance), data=arabica) +
  geom_smooth()

ggplot(aes(x=total_cup_points, y=clean_cup), data=arabica) +
  geom_smooth()

#Look at distribution of DV
ggplot(aes(x=total_cup_points), data=arabica) +
  geom_histogram()

#On the histogram above, we can see that some points are 0, which might mess up our model
arabica <- arabica %>% 
  filter(total_cup_points > 1)

#Check out correlation of predictors
arabica %>% 
  select(aroma, flavor, balance, clean_cup) %>% 
  cor() %>% 
  corrplot(method="number")
```

Then prepare your predictors and fit the most appropriate type of model: 

```{r, error=FALSE, warning=FALSE}
arabica <- arabica %>% 
  mutate(aroma_z = scale(aroma, center = TRUE, scale = TRUE),
         flavor_z = scale(flavor, center = TRUE, scale = TRUE),
         balance_z = scale(balance, center = TRUE, scale = TRUE), 
         clean_cup_z = scale(clean_cup, center = TRUE, scale = TRUE))

arabica_model_full <- lm(total_cup_points ~ balance_z + aroma_z + flavor_z + clean_cup_z, data = arabica)

tidy(arabica_model_full) %>% 
  mutate(estimate = round(estimate, 2))

#Collinearity looks fine below but is borderline... Otherwise, could have tried a reduced model
arabica_model_1 <- lm(total_cup_points ~ aroma_z + clean_cup_z, data = arabica)
arabica_model_2 <- lm(total_cup_points ~ flavor_z + clean_cup_z, data = arabica)
arabica_model_3 <- lm(total_cup_points ~ balance_z + clean_cup_z, data = arabica)

AIC(arabica_model_1, arabica_model_2, arabica_model_3)
```

And check the assumptions:
```{r, error=FALSE, warning=FALSE}
## Collinearity (should be under 10, or under 3/4 if you want to be more strict)
vif(arabica_model_full) 

## Check distribution of residuals a la Winter 2020 p110
res <- residuals(arabica_model_full)
par(mfrow = c(1,3))

###histogram looks ok
hist(res)

###QQ plot 
qqnorm(res)

###Resids look ok, but looks like removing outliers might have improved our fit (i.e. the 65 score)
plot(fitted(arabica_model_full), res)
```

Check model fit. If appropriate, use predict() to see what the model predicts for different values: 
```{r, error=FALSE, warning=FALSE}
#(1)
## Measures of fit (r-squared, etc.)
glance(arabica_model_full)

## Predict new values
arabica_model_raw <- lm(total_cup_points ~ balance + aroma + flavor + clean_cup, data = arabica)
predictions <- tibble(balance = seq(0,10), flavor = seq(0,10), aroma = seq(0,10), clean_cup = seq(0,10))
my_coffee <- tibble(balance = 6, flavor = 8, aroma = 9, clean_cup = 4)
my_coffee$predicted_cupscore <- round(predict(arabica_model_raw, my_coffee), 0)
my_coffee
```


#### Question 2

First, do some exploratory data analysis about your question. This may include some descriptive statistics, some graphing, (tidyverse) summarizing, checking out correlations, etc.:

```{r, error=FALSE, warning=FALSE}
#Plot graphs of predictors vs. DV

coffee %>% ggplot() +
  aes(x = species, y = acidity) + 
  geom_violin()

coffee %>% ggplot() +
  aes(x = species, y = sweetness) + 
  geom_violin()

coffee %>% ggplot() +
  aes(x = species, y = clean_cup) + 
  geom_violin()

coffee %>% ggplot() +
  aes(x = species, y = balance) + 
  geom_violin()

#In the violin plot, we can see that some points are 0, which might mess up our model
coffee <- coffee %>% 
  filter(aroma > 6, clean_cup > 6)
```

Check correlations of predictors
```{r, error=FALSE, warning=FALSE}
coffee %>% 
  select(acidity, sweetness, clean_cup, balance) %>% 
  cor() %>% 
  corrplot(method="number")
```
Balance and acidity are correlated, but logistic regressions can handle some colinearity.

Centering and scaling predictors:
```{r, error=FALSE, warning=FALSE}
coffee <- coffee %>% 
  mutate(acidity_z = scale(acidity, center = TRUE, scale = TRUE),
         sweetness_z = scale(sweetness, center = TRUE, scale = TRUE),
         balance_z = scale(balance, center = TRUE, scale = TRUE), 
         clean_cup_z = scale(clean_cup, center = TRUE, scale = TRUE))
```

Specifying model:
```{r, error=FALSE, warning=FALSE}
species_model_1 <- glm(species ~ sweetness_z + clean_cup_z, 
                       data = coffee, family = binomial)
species_model_2 <- glm(species ~ acidity_z + sweetness_z + clean_cup_z, 
                       data = coffee, family = binomial)
species_model_3 <- glm(species ~ acidity_z + sweetness_z + clean_cup_z + balance_z,
                       data = coffee, family = binomial)

anova(species_model_1, species_model_2, species_model_3, test = "Chisq")

species_lrm_model <- lrm(species ~ sweetness_z + clean_cup_z,
                       data = coffee)

summary(species_model_1)
confint(species_model_1)

species_lrm_model
```

And check the assumptions:
```{r, error=FALSE, warning=FALSE}
rms::vif(species_lrm_model)

car::influencePlot(species_model_1)
coffee[c(716, 1055, 1250, 1283, 1322, 1323),]
```

Check model fit. Use predict() to see the performance: 
```{r, error=FALSE, warning=FALSE}
probability.table <- fitted(species_model_1)
head(probability.table)

coffee$predicted <- predict(species_model_1, newdata = coffee)
coffee <- coffee %>% 
  mutate(pred_cat = if_else(predicted > 0, "Robusta", "Arabica"))

coffee %>% select(species, predicted, pred_cat)
coffee %>% filter(species != pred_cat) %>% select(species, pred_cat)

ctable <- table(coffee$species, coffee$pred_cat)
ctable
round((sum(diag(ctable))/sum(ctable))*100,2)
```


### Graphing

For your hypothesis above, graph the model output with a dot-whisker plot or other appropriate plot style. Use the link below and Ch 11 in Winter 2020:
https://cran.r-project.org/web/packages/dotwhisker/vignettes/dotwhisker-vignette.html

```{r, error=FALSE, warning=FALSE}
#(1)
mycoefs <- arabica_model_full %>% 
  tidy(conf.int = TRUE) %>% 
  filter(term != "(Intercept)")

mycoefs %>% 
  ggplot(aes(x=term, y=estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width = 0.2) +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip()


#(2)
dwplot(list(species_model_1, species_model_2, species_model_3)) +
  theme_light()

dwplot(list(species_model_1, species_model_2, species_model_3),
       vline = geom_vline(xintercept = 0, colour = "grey60", linetype = 2)) %>%
    relabel_predictors(c(acidity_z = "Acidity rating",                       
                         sweetness_z = "Sweetness rating", 
                         clean_cup_z = "Clean cup rating", 
                         balance_z = "Balance")) +
     theme_bw() + xlab("Coefficient Estimate") + ylab("") +
     geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
     ggtitle("Predicting type of coffee bean") +
     theme(plot.title = element_text(face="bold"),
           legend.justification = c(0, 0), 
           legend.background = element_rect(colour="grey80"),
           legend.title = element_blank()) 
```

### Reporting

Now imagine that you want to present your work above. Write your analysis using precise and accurate statistical terminology. 

...

If you have time, make a nice markdown file with the important output, graphs and interpretation (without showing any unneccesary code.)

